---
title: "MC_achilles"
output:
  html_document: default
  pdf_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

load("./Data/cellline_datasets.rda")
load("./Data/prism_datasets.rda")

```

```{r}
library("ggpubr")
```


4 prism.achilles

4.1 Cleanup of prism.achilles 

4.1.1 na removal

These three Rows had to be removed because 
Three rows of our prism.achilles dataframe contain NAs. These are the corresponding DepMap IDs
```{r}
# vector with NA rows - na.achilles
na.achilles <- c(277,309,323)
# Depmap ids of NA containing cellines
rownames(prism.achilles[na.achilles,])
# removing NA containing rows from achilles
prism.achilles <- prism.achilles[-na.achilles,]

```

We make Prism.achilles numeric

```{r}
prism.achilles.numeric <- apply(prism.achilles, 2, as.numeric)
# step below conserves rownames 
rownames(prism.achilles.numeric) <- rownames(prism.achilles)
```

10.Intermission - Sort Prism.cl by depmap_id resulting in prism.cl.ordered

```{r}
# order prism.cl by DepMap_ID resulting in prism.cl.ordered
prism.cl.ordered <- prism.cl[order(prism.cl$DepMap_ID),]
# rownames are scrambled now -> reset
rownames(prism.cl.ordered) <- 1:nrow(prism.cl.ordered)
```

```{r}
# vector a with length of nrow prism.cl
a <- c(1:nrow(prism.cl.ordered))
# because prism cl is ordered we can easily remove the cellines, that have no achilles value, from prism.cl.ordered
# This loop checks if a certain celline has an achilles value and correspondingly writes a 0 or 1 into vector a
for(n in 1:481) {a[n] <- (prism.cl.ordered[n,1] %in% rownames(prism.achilles))}
# vector b contains the prism.cl rows / cellines that do not have an achilles value
b <- which(a == 0)
# remove prism.cl rows that don't have a corresponding achilles value
prism.cl.ordered.achilles <- prism.cl.ordered[-b, ]
# check dimensions of prism.cl.ordered.achilles
# reset rownames of prism.cl.ordered.achilles
rownames(prism.cl.ordered.achilles) <- 1:nrow(prism.cl.ordered.achilles)
```
Find PanCan Cellines in prism.cl.achilles
```{r}
# pancan.rownumbers is a vector with the row numbers that are pancreatic cancer cellines
pancan.rownumbers <- c(which(prism.cl.ordered.achilles[,20] == 'Pancreatic Cancer'))
# pancan.cl.names contains the DepMap IDs of the pancan cellines
pancan.cl.names <- prism.cl.ordered.achilles[pancan.rownumbers,1]
```


The data frame prism.achilles consists of gene knockdown scores. The score is a measure of how essential/important is a particular gene for the cell survival. This score reflects whether upon knocking down that genes does the cell reduce its proliferation or increases it or has no change. Smaller values refers to higher essentiality. The rows of this matrix are the gene names and the columns are the cancer cell line identifiers.

we Exploration of prism.achilles.numeric

```{r}
# More exploration

prism.achilles.numeric.pancan <- prism.achilles.numeric[pancan.rownumbers,]
```
We confirm that Our prism.achilles.numeric and prism.achilles.numeric.pancan dataframes are numeric and can be used for plotting histograms or calculating quantiles.
```{r}
rownames(prism.achilles.numeric) <- rownames(prism.achilles)
```
New Dataframe for Analysis of prism.achilles.numeric
```{r}
# creates Dataframe with 10 rows and the same column count and names as prism.achilles
achilles.research.dataframe <- prism.achilles[1:10,]
# this sets all values of the research.dataframe to 0
achilles.research.dataframe[1:10,1:18119] <- 0
# this sets all rownames according to their future contents
rownames(achilles.research.dataframe) <- c('max','mean','min','10%quantile','Pancanmean','is(r5<r4)','stdev','empty8','empty9','empty10')

# fills first row of research.dataframe with the maximum value of the according prism.achilles.numeric column
achilles.research.dataframe[1,] <- apply(prism.achilles.numeric, 2, max)
# means of prism.achilles.numeric
achilles.research.dataframe[2,] <- apply(prism.achilles.numeric, 2, mean)
# min of prism.achilles.numeric
achilles.research.dataframe[3,] <- apply(prism.achilles.numeric, 2, min)
# 10% quantile of prism.achilles.numeric
achilles.research.dataframe[4,] <- apply(prism.achilles.numeric, 2, function(x){quantile(x,prob=0.1)})
# mean of prism.achilles.numeric.pancan
achilles.research.dataframe[5,] <- apply(prism.achilles.numeric.pancan, 2, mean)
# is 1 if mean of prism.achilles.numeric.pancan is smaller that 10% quantile of prism.achilles.numeric
achilles.research.dataframe[6,] <- apply(achilles.research.dataframe, 2, function(x) {return(x[4] > x[5])})
# var of prism.achilles.numeric
achilles.research.dataframe[7,] <- apply(prism.achilles.numeric, 2, var)

```

The four interesting genes
```{r}

achilles.research.dataframe[1:7,which(achilles.research.dataframe[6,] == 1)]
```
Shapiro tests of the four interesting prism.achilles columns shows that their values are not normally distributed. Thus the comparison of the prism.achilles.numeric.pancan mean and prism.achilles.numeric 10% quantile are of no importance until further examination. Shown is the result of a Shapiro-Wilk normality test on column 4283 of our prism.achilles.numeric dataset.
```{r}
# these values show that the interesting columns are not completely normally distributed
shapiro.test(prism.achilles.numeric[,4283])
```
A qqplot between prism.achilles.numeric[,4283] and an according normal distribution results in the following graph. This shows that the values do not correspond to one single normal distribution.
```{r}
ggqqplot(prism.achilles.numeric[,4283])
```




k-means clustering

Using the most variable, thus informative genes
We first reduce our dataset to take the most variable gene, which are expected to carry most of the information about the samples:
```{r}
## compute variance over all rows of the data.frame (hence genes)
topVar = apply(prism.achilles.numeric, 2, var) 
```
```{r variance filtering for clustering}
## filtering to keep genes which have a variance greater than the 75% percentile
prism.achilles.numeric.topVar = prism.achilles.numeric[,topVar > quantile(topVar, probs = 0.75)] 
```



