---
title: "MC_achilles"
output:
  html_document: default
  pdf_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

load("./Data/cellline_datasets.rda")
load("./Data/prism_datasets.rda")
library("ggpubr")
```

4 prism.achilles

4.1 Cleanup of prism.achilles 

4.1.1 na removal

```{r find NA}
# Where are our NAs?
which(is.na(prism.achilles[,14]))
prism.achilles[275:280,14]
prism.achilles[320:325,14]
```

```{r}
# vector with NA rows - na.achilles
na.achilles <- c(277,309,323)
# Depmap ids of NA containing cellines
rownames(prism.achilles[na.achilles,])
# how many total NAs?
sum(is.na(prism.achilles[na.achilles,]))
# removing NA containing rows from achilles
prism.achilles <- prism.achilles[-na.achilles,]
# checking if any NAs left
sum(is.na(prism.achilles))
# checking dim of prism.achilles - no genes removed - only NA cell lines removed
dim(prism.achilles)
```

4.1.2 Make Prism.achilles numeric

```{r}
prism.achilles.numeric <- apply(prism.achilles, 2, as.numeric)
# step below conserves rownames 
rownames(prism.achilles.numeric) <- rownames(prism.achilles)
```

10.Intermission - Sort Prism.cl by depmap_id resulting in prism.cl.ordered

```{r}
# show slice of prism.cl - rows are not yet ordered
prism.cl[1:3,1:3]
# order prism.cl by DepMap_ID resulting in prism.cl.ordered
prism.cl.ordered <- prism.cl[order(prism.cl$DepMap_ID),]
# dimmension check - hasn't changed
dim(prism.cl.ordered)
# rownames are scrambled now -> reset
rownames(prism.cl.ordered) <- 1:nrow(prism.cl.ordered)
# check rowname order
rownames(prism.cl.ordered)[1:30]
# check slice - everything worked
prism.cl.ordered[1:30,1:3]
```

```{r}
# vector a with length of nrow prism.cl
a <- c(1:nrow(prism.cl.ordered))
# because prism cl is ordered we can easily remove the cellines, that have no achilles value, from prism.cl.ordered
# This loop checks if a certain celline has an achilles value and correspondingly writes a 0 or 1 into vector a
for(n in 1:481) {a[n] <- (prism.cl.ordered[n,1] %in% rownames(prism.achilles))}
# vector b contains the prism.cl rows / cellines that do not have an achilles value
b <- which(a == 0)
# remove prism.cl rows that don't have a corresponding achilles value
prism.cl.ordered.achilles <- prism.cl.ordered[-b, ]
# check dimensions of prism.cl.ordered.achilles
dim(prism.cl.ordered.achilles)
# reset rownames of prism.cl.ordered.achilles
rownames(prism.cl.ordered.achilles) <- 1:nrow(prism.cl.ordered.achilles)
# compare rownames of achilles with cl - checking for same order and same elements - they are equal
prism.cl.ordered.achilles[1:5,1]
rownames(prism.achilles.numeric)[1:5]
```
Find PanCan Cellines in prism.cl.achilles
```{r}
# pancan.rownumbers is a vector with the row numbers that are pancreatic cancer cellines
pancan.rownumbers <- c(which(prism.cl.ordered.achilles[,20] == 'Pancreatic Cancer'))
# pancan.cl.names contains the DepMap IDs of the pancan cellines
pancan.cl.names <- prism.cl.ordered.achilles[pancan.rownumbers,1]
# checking if this was done correctly
prism.cl.ordered.achilles[pancan.rownumbers,20]
# this checks if all steps before were done correctly - row numbers of cl and achilles correspond to the same depmap ids
rownames(prism.achilles.numeric)[pancan.rownumbers]
pancan.cl.names
```

```{r}
# More tests if everything is ordered correctly
rownames(prism.cl.ordered.achilles)[1:10]
prism.cl.ordered.achilles[1:10,1]
rownames(prism.achilles.numeric)[1:10]
```


The data frame prism.achilles consists of gene knockdown scores. The score is a measure of how essential/important is a particular gene for the cell survival. This score reflects whether upon knocking down that genes does the cell reduce its proliferation or increases it or has no change. Smaller values refers to higher essentiality. The rows of this matrix are the gene names and the columns are the cancer cell line identifiers.

Exploration of prism.achilles.numeric
```{r}
dim(prism.achilles.numeric)
prism.achilles.numeric[1:10,1:2]
```
```{r}
# More exploration
pancan.rownumbers
prism.achilles.numeric.pancan <- prism.achilles.numeric[pancan.rownumbers,]
prism.achilles.numeric.pancan[1:10,1]
```
```{r}
# just confirming it is still numeric
is.numeric(prism.achilles.numeric.pancan)
hist(prism.achilles.numeric.pancan[10,])
hist(prism.achilles.numeric[1,])
quantile(prism.achilles.numeric.pancan,probs=seq(0,1,.1))
```
```{r}
# checking that i can apply the quantile function...
quantile(prism.achilles.numeric,probs=seq(0,1,.1))
quantile(prism.achilles.numeric,prob=0.1)
```
```{r}
prism.achilles.numeric[345:347,1:5]
dim(prism.achilles.numeric)
dim(prism.achilles)
rownames(prism.achilles.numeric) <- rownames(prism.achilles)
```
New Dataframe for Analysis of prism.achilles.numeric
```{r}
# creates Dataframe with 10 rows and the same column count and names as prism.achilles
achilles.research.dataframe <- prism.achilles[1:10,]
# this sets all values of the research.dataframe to 0
achilles.research.dataframe[1:10,1:18119] <- 0
# this sets all rownames according to their future contents
rownames(achilles.research.dataframe) <- c('max','mean','min','10%quantile','Pancanmean','is(r5<r4)','stdev','empty8','empty9','empty10')

# fills first row of research.dataframe with the maximum value of the according prism.achilles.numeric column
achilles.research.dataframe[1,] <- apply(prism.achilles.numeric, 2, max)
# means of prism.achilles.numeric
achilles.research.dataframe[2,] <- apply(prism.achilles.numeric, 2, mean)
# min of prism.achilles.numeric
achilles.research.dataframe[3,] <- apply(prism.achilles.numeric, 2, min)
# 10% quantile of prism.achilles.numeric
achilles.research.dataframe[4,] <- apply(prism.achilles.numeric, 2, function(x){quantile(x,prob=0.1)})
# mean of prism.achilles.numeric.pancan
achilles.research.dataframe[5,] <- apply(prism.achilles.numeric.pancan, 2, mean)
# is 1 if mean of prism.achilles.numeric.pancan is smaller that 10% quantile of prism.achilles.numeric
achilles.research.dataframe[6,] <- apply(achilles.research.dataframe, 2, function(x) {return(x[4] > x[5])})
# var of prism.achilles.numeric
achilles.research.dataframe[7,] <- apply(prism.achilles.numeric, 2, var)

```


```{r}
achilles.research.dataframe[1:7,1]
which(achilles.research.dataframe[6,] == 1)
achilles.research.dataframe[1:7,which(achilles.research.dataframe[6,] == 1)]
```
```{r}
# these values show that the interesting columns are not completely normally distributed
shapiro.test(prism.achilles.numeric[,4283])
shapiro.test(prism.achilles.numeric[,7638])
shapiro.test(prism.achilles.numeric[,7955])
shapiro.test(prism.achilles.numeric[,17229])
```

ggqqplot(prism.achilles.numeric[,4283])
ggqqplot(prism.achilles.numeric[,7638])
ggqqplot(prism.achilles.numeric[,7955])
ggqqplot(prism.achilles.numeric[,17229])




merlinsvariable <- 10000
cor.mat.achilles <- cor(prism.achilles.numeric.pancan[,1:merlinsvariable])
cor.mat.achilles[779,782]
interesting.cor.values <- which(cor.mat.achilles > 0.90 & cor.mat.achilles < 1 ) 
interesting.cor.rows <- (((interesting.cor.values - (interesting.cor.values %% merlinsvariable)) /merlinsvariable) +1)
interesting.cor.cols <- (interesting.cor.values %% merlinsvariable)
cor.mat.achilles[interesting.cor.rows,  interesting.cor.cols]



k-means clustering

Using the most variable, thus informative genes
We first reduce our dataset to take the most variable gene, which are expected to carry most of the information about the samples:
```{r}
## compute variance over all rows of the data.frame (hence genes)
topVar = apply(prism.achilles.numeric, 2, var) 
summary(topVar)
```
```{r variance filtering for clustering}
## filtering to keep genes which have a variance greater than the 75% percentile
prism.achilles.numeric.topVar = prism.achilles.numeric[,topVar > quantile(topVar, probs = 0.75)] 
dim(prism.achilles.numeric.topVar)
```


Clustering by depmap
```{r}
mydata <- prism.achilles.numeric.topVar
# Determine number of clusters
wss.depmap <- (nrow(mydata)-1)*sum(apply(mydata,2,var))
wss.depmap
for (i in 2:100) wss.depmap[i] <- sum(kmeans(mydata, centers=i)$withinss)
```

```{r}
wss.depmap
plot(1:100, wss.depmap, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")
plot(1:20, wss.depmap[1:20], type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")
```


```{r}
km.depmap = kmeans(x=prism.achilles.numeric.topVar, 
            centers = 2, 
            nstart = 10)
km.depmap$cluster[1:10]

```


```{r}
summary(km.depmap)
pancan.cl.names
km.depmap$cluster[pancan.cl.names]
```

Clustering by genes
```{r}
mydata <- prism.achilles.numeric.topVar
# Determine number of clusters

wss.genes <- (ncol(mydata)-1)*sum(apply(mydata,1,var))
dim(mydata)
for (i in 500:501) wss.genes[i] <- sum(kmeans(t(mydata), centers=i, iter.max = 500)$withinss)
```



wss.genes
dim(wss.genes)
plot(wss.genes[100:1000], type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")
plot(wss.genes[100:500], type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")



```{r}
km.genes = kmeans(x=prism.achilles.numeric.topVar, 
            centers = 2, 
            nstart = 10)
km.genes$cluster[1:10]

```


```{r}
summary(km.genes)
pancan.cl.names
km.genes$cluster[pancan.cl.names]
```
